CSE530 Project Final Report\\ Bandwidth-Aware Memory Hierarchy Design with Hybrid Memory Technologies

### 2010-11-18 paper.tex
abstract

\begin{abstract}
\boldmath

@@

Many of modern chip-multiprocessors are designed to perform well on various
applications to achieve high performance by exploiting their inherent
parallelism. Such systems support large number of threads and single instruction
multiple data~(SIMD) execution, which puts a lot of pressure on the memory
system. Memory latency is typically not a bottleneck, since the latency can be
hidden via multi-threading or hardware prefetching. To this end, bandwidth becomes a
potential bottleneck. A high rate of computing often brings in a high rate of
data transitions. In some cases, the working set of an application fits in the
on-die caches, which can typically provide sufficient bandwidth to keep up with
the processing cores. However, if the working set does not fit in the on-die
caches, the main memory needs to provide much of the data. Since the bandwidth
of off-chip main memory is quite limited, applications with such working sets
are potentially bandwidth-bound. Therefore, it is crucial to design the memory
hierarchy to overcome the bandwidth limitation.

@@

 We will also use
Simics~\cite{Magnusson:2002:simics-orig-paper} as the simulator to evaluate our
design method. The benchmarks will be selected from bandwidth-bound
applications, which possibly include benchmarks from SPEC
CPU2006~\cite{SPEC:2006} and NPB~\cite{NPB}.\vspace{0.15in}

@@

\begin{table}[t]
  \centering
  \caption{Baseline CMP configuration.}
  \vspace{0.1in}
  \begin{tabular}{l|c}
    \hline
    \multicolumn{2}{|l|}{\textbf{Core}}\\
    \hline
    No. of cores & 8\\
    \hline
    Frequency & 1GHz\\
    \hline
    Core architecture & in-order, 14-stage pipline, 2 ALUs, 2 FPUs\\
    \hline
    \multicolumn{2}{|l|}{\textbf{Memory}}\\
    \hline
    \multirow{2}{*}{Private caches} & L1 I-cache: SRAM, 8 x 64KB, 64B line, 2-way, write-through \\
                                    & L1 D-cache: SRAM, 8 x 64KB, 64B line, 2-way, write-through \\
    \hline
    \multirow{2}{*}{Shared caches} & Case 1: Pure SRAM, 64B line, 8-way, write-back \\
                                   & Case 2: Hybrid~(SRAM, MRAM, RRAM, eDRAM), 64B line, 8-way, write-back\\
    \hline
    Main memory & 4GB\\
    \hline
  \end{tabular}
  \label{table:cmp-config}
\end{table}

The benchmarks are selected from PARSEC benchmark suite~\cite{PARSEC:2008} with
multithreaded programs, which focus on emerging workloads that are designed to
be representative of next-generation shared-memory programs for CMPs. Since the
performance of different memory technologies are closely related to read and
write intensities, we selected some workloads that vary in the average numbers
of L2 cache read per thousand instructions~(RPKI) and write per thousand
instructions~(WPKI), which are listed in Table~\ref{table:benchmarks}.

\begin{table}[htbp!]
  \centering
  \caption{Characteristics of selected benchmarks.}
  \vspace{0.1in}
  \begin{tabular}{l|c|c||l|c|c}
    \hline
    \textbf{Benchmarks} & \textbf{RPKI} & \textbf{WPKI} &
    \textbf{Benchmarks} & \textbf{RPKI} & \textbf{WPKI}\\
    \hline
    blackscholes & 22.8 & 61.7 & ferret & 5.7 & 173.4\\
    \hline
    bodytrack & 5.4 & 139.5 & fluidanimate & 2.6 & 70.1\\
    \hline
    canneal & 5.4 & 29.3 & streamcluster & 17.3 & 16.9\\
    \hline
    facesim & 6.0 & 102.4 & swaptions & 2.6 & 121.4\\
    \hline
  \end{tabular}
  \label{table:benchmarks}
\end{table}

>>>>

\begin{abstract}
  \boldmath Chip-multiprocessor~(CMP) is a promising solution for high
  performance computing systems. However, the limited increases in main memory
  bandwidth results in a growing bandwidth gap between the processor cores and
  the main memory, which becomes a potential obstacles to system performance
  improvement. In this project, we explore bandwidth-aware memory hierarchy
  design with various emerging memory technologies. Based on the observations
  obtained from our pre-exploration of various emerging memory technologies, we
  evaluate systerm performance obtained with various cache hierarchy
  configurations. With different applications, we evaluate the optimal shared
  cache configurations in a CMP system, in terms of the number of shared cache
  levels, memory technology and capacity of each level. We show that cache
  hierarchy with hybrid memory technologies benefits system performance with
  non-write-intensive applications.

@@

Many of modern chip-multiprocessors are designed to perform well on various
applications to achieve high performance by exploiting their inherent
parallelism. Such systems support large number of threads and single instruction
multiple data~(SIMD) execution, which puts a lot of pressure on the memory
system. Typically, memory latency is not a bottleneck, since the latency can be
hidden via multi-threading or hardware prefetching. To this end, bandwidth
becomes a potential bottleneck. A high rate of computing often brings in a high
rate of data transitions. In some cases, the working set of an application fits
in the on-die caches, which can typically provide sufficient bandwidth to keep
up with the processing cores. However, if the working set does not fit in the
on-die caches, the main memory needs to provide much of the data. Since the
bandwidth of off-chip main memory is quite limited, applications with such
working sets are potentially bandwidth-bound. Therefore, it is crucial to design
the memory hierarchy to overcome the bandwidth limitation.

@@

Based on the parameters of different cache configurations collected from our
modified version of CACTI~\cite{CACTI}, we evaluate both pure SRAM-based and
hybrid cache hierarchy designs. In this section, we show how bandwidth-aware
memory hierarchy design method help with improving the system performance.\vspace{0.15in}

@@

\begin{table}[t]
  \centering
  \caption{Baseline CMP configuration.}
  \vspace{0.1in}
  \begin{tabular}{l|c}
    \hline
    \multicolumn{2}{l}{\textbf{Core}}\\
    \hline
    No. of cores & 8\\
    \hline
    Frequency & 1GHz\\
    \hline
    Core architecture & in-order, 14-stage pipline, 2 ALUs, 2 FPUs\\
    \hline
    \multicolumn{2}{l}{\textbf{Memory}}\\
    \hline
    \multirow{2}{*}{Private caches} & L1 I-cache: SRAM, 8 x 64KB, 64B line, 2-way, write-through \\
                                    & L1 D-cache: SRAM, 8 x 64KB, 64B line, 2-way, write-through \\
    \hline
    \multirow{2}{*}{Shared caches} & Case 1: Pure SRAM, 64B line, 8-way, write-back \\
                                   & Case 2: Hybrid~(SRAM, MRAM, RRAM, eDRAM), 64B line, 8-way, write-back\\
    \hline
    Main memory & 4GB\\
    \hline
  \end{tabular}
  \label{table:cmp-config}
\end{table}

The benchmarks are selected from PARSEC benchmark suite~\cite{PARSEC:2008} with
multithreaded programs, which focus on emerging workloads that are designed to
be representative of next-generation shared-memory programs for CMPs. Since the
performance of different memory technologies are closely related to read and
write intensities, we selected some workloads that vary in the average numbers
of L2 cache read per thousand instructions~(RPKI) and write per thousand
instructions~(WPKI), which are listed in
Table~\ref{table:benchmarks}.\vspace{0.3in}

\begin{table}[t]
  \centering
  \caption{Characteristics of selected benchmarks.}
  \vspace{0.1in}
  \begin{tabular}{l|c|c||l|c|c}
    \hline
    \textbf{Benchmarks} & \textbf{RPKI} & \textbf{WPKI} &
    \textbf{Benchmarks} & \textbf{RPKI} & \textbf{WPKI}\\
    \hline
    blackscholes & 22.8 & 61.7 & ferret & 5.7 & 173.4\\
    \hline
    bodytrack & 5.4 & 139.5 & fluidanimate & 2.6 & 70.1\\
    \hline
    canneal & 5.4 & 29.3 & streamcluster & 17.3 & 16.9\\
    \hline
    facesim & 6.0 & 102.4 & swaptions & 2.6 & 121.4\\
    \hline
  \end{tabular}
  \label{table:benchmarks}
\vspace{0.5in}
\end{table}
